### 2021-04-28 Minutes Modelleren-B Week 4.2.tex

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Notule Week 4.2}
\author{Modelleren-B groep 15}
\date{April 2021}

\begin{document}
Functies maken:
-Functie van Excel naar Python omzetten
-Cost function

Ideeën:
-
\end{document}