### 2021-05-25 ILP.tex
Add files via upload
Eerste ILP formulatie

\documentclass[11pt]{article} % use larger type; default would be 10pt
\usepackage{amsmath}
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....

\title{ILP of Parcel Delivery}
\author{Bart Stolk}

\begin{document}
\maketitle

\section{Input}
\begin{align*}
c_{ij} &= \text{cost between city i and j}  \quad i,j=1,\ldots,N \\
w_{ij} &= \text{flow between city i and j} \quad i,j=1,\ldots,N \\
f_{i} &= \text{cost to turn city i into a hub} \quad i=1,\ldots,N \\
\chi &= \text{collection cost multiplier} \\
\alpha &= \text{transfer cost multiplier} \\
\delta &= \text{distribution cost multiplier} \\
\end{align*}

\section{Variables}
\begin{align*}
h_{i} &=
\begin{cases} 
1, & \text{if city i is a hub} \quad i=1,2,\ldots,N \\ 
0, & \text{otherwise} 
\end{cases} \\ 
e_{ij} &=
\begin{cases} 
1, & \text{if there is and edge between city i and j} \quad i,j=1,2,\ldots,N \\ 
0, & \text{otherwise} 
\end{cases} \\
n_{ij} &= 
\begin{cases} 
0, & \text{if } h_{i} = 1 \text{ and } h_{j}=1 \quad i,j=1,2,\ldots,N \\ 
e_{ij}, & \text{otherwise}
\end{cases}
\end{align*}

\section{Objective function}
$$min\,z=\sum^N_{i=1}\sum^N_{j=1}\sum^N_{k=1}\sum^N_{l=1}(w_{ij}(\chi c_{ik}e_{ik}+(\alpha c_{kl}+\delta c_{lj})e_{lj})) + \sum^N_{i=1}f_{i}h_{i}$$
s.t.
\begin{align*}
e_{ij}-h_{i}-h_{j} &= 0 \quad i,j=1,2,\ldots,N \text{ for } i \ne j \\
e_{ii}-h_{i} &= 0 \quad i=1,2,\ldots,N \\
\sum^N_{j=i}n_{ij} &= 1 \quad i=1,2,\ldots,N \\
\sum^N_{i=1}h_{i} &\geq 1 \quad i=1,2,\ldots,N
\end{align*}
\end{document}