### 2021-05-27 draft_report.tex
draft

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{authblk}

\title{Draft Report Modelleren-B}
\author{Bart Stolk \and Dyami Silvério \and Wessel van Sommeren \and Severijn Wiechers}
\date{03 June 2021}

\begin{document}

\maketitle

\newpage
\section*{Summary}
\addcontentsline{toc}{section}{Summary}

\newpage
\tableofcontents

\newpage
\section{Introduction}
Reducing costs is a valuable tool for any package delivery cooperation to stay relevant and make profits. This report will try to show how it is possible to create an algorithm that gives an optimum 
\newpage
\section{Intuitive Algorithm}

    \subsection{Calculating the cost for multiple hubs}
    \subsection{The different algorithms used}
    \subsection{Result of the combined algorithms}

\newpage  
\section{The ILP solver}
We are given a set N of nodes where we can locate a hub. The cost of opening a hub at node i $\in$ N is $F_i$ and the cost of transporting a parcel to node j from node i is $c_{i_j}$, i $\in$ N, j $\in$ N. We need to decide on which nodes we shall open a hub, and from which hub the nodes will be connected. The goal is to minimize the total cost. 
    \subsection{ILP implementation}
    \subsection{IlP result}

\newpage
\section{Expanded intuitive algorithm}

\newpage
\section{Conclusion}

\section{Appendix} 
\end{document}