### 2021-06-01 Minutes/Minutes Modelleren-B week 4.6 .tex

\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Minutes Modelleren 25-05-2021}
\author{Severijn Wiechers}
\date{May 2021}

\begin{document}

\maketitle
25-05-2021
\section{Attendance}
\begin{itemize}
    \item Wessel was still in the train so he couldn't be in the meeting
    \item Dyami had a doctors appointment but was in the meeting for the majority of the first hour
    \item Bart and Severijn were present the entire meeting
\end{itemize}

\section{what do we need for the draft report}
\begin{itemize}
    \item The complete intuitive algorithm. Here we will write down all the made algorithms and the conclusion of the chosen hubs.
    \item The answer to the problem via the ilp solver

\end{itemize}

\section{What have we done today}
\begin{itemize}
    \item By combining the results of all the intuitive algorithms we think that the combination of the hubs [6, 14, 8] will have one of the lowest possible costs at 144532 euros in total
    \item We still need to get an answer out of the ILP solver, Bart will try to do that today
\end{itemize}

\section{What do we still have to do}
\begin{itemize}
    \item we still have to let the cost function show the different cities that are connected per hub
\end{itemize}

\newpage
27-05-2021
\section{Attendance}
\begin{itemize}
    \item Everyone was were present the entire meeting
\end{itemize}

\section{Draft Report}
The entire layout of the draft report is made. Furthermore, the first two chapters are almost finished. Before the next meeting the goal is to finish the chapters about the intuitive algorithm. Next Tuesday the final chapters of the report will be written.

\section{What have we done today}
\begin{itemize}
    \item We have written the first chapters of the draft report
    \item We have discussed the peer review; there weren't any explicit thing that stood out // that we wanted to discuss further

\end{itemize}

\section{What do we still have to do}
\begin{itemize}
    \item Bart will finish the ILP solver before the next meeting
    \item Every member will finish their part in the report explaining their intuitive algorithm.
\end{itemize}
\end{document}

Hoe verhoud de ilp tot intuitief
crititische blik op de uitkomsten van het algoritme