### 2021-06-03 ILP/ILP.tex
Update ILP.tex

\documentclass[11pt]{article} % use larger type; default would be 10pt
\usepackage{amsmath}
\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....

\title{ILP of Parcel Delivery}
\author{Bart Stolk}

\begin{document}
\maketitle

\subsection*{Input}
\begin{align*}
c_{ij} &= \text{cost between city i and j}  &i,j \in N\\
w_{ij} &= \text{flow between city i and j} &i,j \in N\\
f_{i} &= \text{cost to turn city i into a hub} &i \in N\\
\chi &= \text{collection cost multiplier} \\
\alpha &= \text{transfer cost multiplier} \\
\delta &= \text{distribution cost multiplier} \\
\end{align*}

\subsection*{Variables}
\begin{align*}
e_{ik} &=
\begin{cases} 
1, & \text{if city i is assigned to hub k} \quad i=1,2,\ldots,N \quad j=1,2,\ldots,N  \\ 
0, & \text{otherwise} 
\end{cases} \\
y_{ijkl} &=
\begin{cases} 
1, & \text{if there is a path } i \rightarrow k \rightarrow l \rightarrow j \quad i=1,2,\ldots,N \quad j=1,2,\ldots,N  \\ 
0, & \text{otherwise} 
\end{cases} \\
\end{align*}

\subsection*{Objective function}
$$min\,z=\sum^N_{i=1}\sum^N_{j=1}\sum^N_{k=1}\sum^N_{l=1}(w_{ij}(\chi c_{ik}+\alpha c_{kl}+\delta c_{lj})y_{ijlk} + \sum^N_{k=1}f_{k}e_{kk}$$

\subsection*{Constraints}
\begin{align*}
\sum^N_{k=1}e_{ik} &= 1 &i \in N \\
e_{ik} & \leq e_{kk} &i,k \in N \\
y_{ijkl} &\leq e_{ik} &i,j,k,l \in N \\
y_{ijkl} &\leq e_{jl} &i,j,k,l \in N \\
y_{ijkl} &\geq e_{ik} + e_{jl} - 1 &i,j,k,l \in N \\
\sum^N_{k=1}f_k e_{kk} &\geq 1
\end{align*}
\end{document}