### 2021-06-08 Minutes/Minutes_Modelleren_4.8.tex
Updates from Overleaf

\documentclass{article}
\usepackage[utf8]{inputenc}

\begin{document}
\section*{Minutes Wednesday}
Aanwezig: Bart, Dyami, Severijn
Voorzitter: Dyami
Secretaris: Bart

\subsection*{Discussed topics}
\begin{itemize}
    \item Incorporating the small data set into the report.
    \item Finalizing the ILP implementation for the expanded problem.
    \item Giving feedback on the draft report of group 5.
    \item Rewriting the ILP definition in the report, put the entire thing at the top
    \item More graphics, pictures and the sorts.
    \item {Pitch, Who and about what}
\end{itemize}

\subsection*{Decisions}
\begin{itemize}
    \item Continue work on the small data set and the ILP implementation after meeting.
    \item Today (Tuesday), we will independently give feedback on portions of the draft report of group 5 and then come together to give feedback.
    \item Severijn is going to present the pitch.
\end{itemize}

\subsection*{Planning}
\begin{itemize}
    \item First draft Pitch done before Thursday.
    \item Finish ILP implementation before Thursday.
\end{itemize}
\end{document}